---
output: pdf_document
---

```{r include=FALSE}
library(formatR)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff = 80), tidy = TRUE)
```

# Intro to Data Science - HW 5

##### Copyright 2023, Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva

```{r}
# Enter your name here: Alanna Pellecchia
```

### Attribution statement: (choose only one and delete the rest)

```{r}
# 1. I did this homework by myself, with help from the book and the professor.
```


```{r}
# load packages
library(RCurl)
library(tidyverse)
library(curl)
```

Reminders of things to practice from previous weeks: <br>
Descriptive statistics: 	mean( ) max( ) min( )<br>
Coerce to numeric:		as.numeric( )<br>

## Part 1: Use the Starter Code 

Below, I have provided a starter file to help you. <br>

Each of these lines of code **must be commented** (the comment must that explains what is going on, so that I know you understand the code and results).

```{r}
#install.packages('RCurl')
#install.packages('jsonlite')
# load and install packages
library(RCurl)
library(jsonlite)
# store url in dataset var
dataset <- getURL("https://intro-datascience.s3.us-east-2.amazonaws.com/role.json")
# convert from json into list 
readlines <- jsonlite::fromJSON(dataset)
# store list data in dataframe 
df <- readlines$objects$person
```

A.	Explore the **df** dataframe (e.g., using head() or whatever you think is best).

```{r}
# use head to see first few rows
head(df)
```

B.	Explain the dataset <br>
o	What is the dataset about? <br>
o	How many rows are there and what does a row represent?<br>
o	How many columns and what does each column represent?<br>

```{r}
# the dataset is a list of senators and their info. 
# there are 100 rows, each row is a senator.
# there are 17 columns.
# there are four ids: bioguideid, osid, pvsid, and cspan id.
# there is first name, last name, middle name, and full name.
# there is also a name mod (ex. Jr) and a nickname.
# there are two gender columns.
# there is the birthday and the link to their site.
# there is also their twitter and youtube username. 
```

## Part 2: Investigate the resulting dataframe 

A.	Describe what you see when you run the **table()** function on the **gender** variable.

```{r}
# it shows the number of female and male entries. 
table(df$gender)
```

A1.	Generate the count of number of females and number of males, using the tidyverse **group_by()**, **summarise()** and **n()** functions.

```{r}
# female is 24, male is 76. 
# the table function above easily generates those numbers. 
```

B. How many senators are women?

```{r}
# 24.
```

C.	How many senators don't have a YouTube account?

```{r}
# check to see how many rows in youtubeid come up as NA
# 27 senators have no account
is.na(df$youtubeid)
```

D. Show how many senators ***do*** have a YouTube account.  

```{r}
# remove NA from df, filter into new dataframe 
youtubers <- filter(df, !is.na(df$youtubeid))
# show # of rows and columns
str(youtubers)
# 73 have a youtube account. 
```

E.	How many women senators have a YouTube account? 

```{r}
# display gender table from youtubers df
table(youtubers$gender)
# 16 women have youtube accounts. 
```

F.	Create a new dataframe called **youtubeWomen** that only includes women senators who have a YouTube account.   

```{r}
# filter for only rows with female as gender
# store in new dataframe
youtubeWomen <- filter(youtubers, youtubers$gender == 'female')
```

G.	What does running this line of code do? Explain in a comment: 

```{r}
# it adds a new column - year
# this is the birth year of each senator in youtubeWomen. 
# it does this by creating a substring from birthday. 
youtubeWomen$year <- substr(youtubeWomen$birthday,1,4)
```

H.	Use this new variable to calculate the mean **birthyear** in **youtubeWomen**. <br>
**Hint:** You may need to convert it to numeric first using the **as.numeric()** function.

```{r}
# convert year from character to numeric
youtubeWomen$year <- as.numeric(youtubeWomen$year)
# take mean = 1954.875
mean(youtubeWomen$year)
```

I.	Make a histogram of the **birthyears** of senators in **youtubeWomen**. Add a comment describing the shape of the distribution.

```{r}
# histogram
hist(youtubeWomen$year)
# normal distribution
```

J. Create a dataframe called **youtubeMen** which only includes male senators with a youTube account. Repeat steps G & H for this dataframe and create a histogram of the birthyears in it. Compare the shape and properties of this histogram to the one in H.

```{r}
# create youtubeMen by filtering youtuber df for male
youtubeMen <- filter(youtubers, youtubers$gender == 'male')
# substring birthday, store in year column
youtubeMen$year <- substr(youtubeMen$birthday,1,4)
# convert year to numeric 
youtubeMen$year <- as.numeric(youtubeMen$year)
# take mean = 1955.439 
mean(youtubeMen$year)
# histogram
hist(youtubeMen$year)
# it is a similar distribution, but a much smoother curve.
```

K. Take a look at [this article](https://www.theguardian.com/us-news/ng-interactive/2018/nov/15/new-congress-us-house-of-representatives-senate) - explore its interactive features and focus specifically on the section on **gender**. Relating what you learned from the article back to our Senate data, who might feel left out and/or unrepresented based on the current gender composition of the Senate? Explain in a brief comment.

```{r}
# anyone who is not a man may feel left out.
# there are far less women senators than men.
# in addition, there appear to be no non-binary senators. 
```

